---
title: "R Notebook"
output: html_notebook
---

link 2 https://www.kaggle.com/datasets/sagarvarandekar/macroeconomic-factors-affecting-us-housing-prices/code


This dataset contains month-wise details about all the macro-economic factors of US over two decades from 05/2002 to 05/2022 which can be used to determine how different factors affect a particular factor.
Sourced and compiled data from official websites of US government like FRED, CENSUS, OECD, Conference Board.
Collected economic data like Income, GDP, different indices, mortgage etc. from FRED official website, demographic data from OECD and consumer confidence Index from Conference Board.

CONSUMER CONF INDEX - Consumer Confidence Index provided by Conference Board
PPI-CONST MAT  - Producers Purchase Index- Construction Materials
CPIALLITEMS - Consumer Price Index- All Items for the US
INFLATION(%) - Inflation rate in the US
MORTGAGE INT. MONTHLY AVG(%) - Average Mortgage interest rate of all Weeks of a particular month
MED HOUSEHOLD INCOME - Median Household Income in the US
CORP. BOND YIELD(%) 
MONTHLY HOME SUPPLY


```{r}
rm(list=ls())
library(tidyverse)
library(lubridate)
library(zoo)
library(skimr)
library(ggcorrplot)
library(lmtest)
library(sandwich)



setwd("C:/Users/49162/OneDrive/Desktop/All_documents/Kaggle/Bank_marketing")
data_us <- read.csv("DATA.csv")

data_us <- tibble(data_us)
```



```{r}
head(data_us)


```

```{r}
colnames(data_us)



```

Renaming the columns
```{r}
names <- str_to_title(colnames(data_us))
names

```


```{r}
new_names <- str_replace_all(names, "[\\.]+", "_")
new_names

new_names1 <- str_replace_all(new_names, c("_$" = "", "^._" = "" ))
new_names1



writeLines(new_names1)

```



```{r}

#new_names <- str_sub(new_names,1,15)
colnames(data_us) <- new_names1
colnames(data_us)

```




```{r}
str(data_us)

```

```{r}


#for(i in seq_along(data_us)){
 # data_us$date[[i]]<- dd_date(data_us$date[[i]])
#}
data_us$Date <- dmy(data_us$Date)
str(data_us)

data_us <- data_us%>% arrange(Date)
head(data_us)

```


```{r}
apply(is.na(data_us), 2, sum)


```
```{r}
tail(data_us, 20)
```



```{r}

attach(data_us)
data_us <- na.locf(data_us, fromLast = FALSE)

apply(is.na(data_us), 2, sum)
```

```{r}
#data_us <- data_us %>% select(!date)

#data_us <- as_tibble(data_us)
#head(data_us)

data_us <- data_us %>% rename(cpi_all_items = Cpiallitems)

library(GrowthRate)

GrowthRate <- function(x){
  y = (x - (dplyr::lag(x)))/dplyr::lag(x)
  return(y*100)
}

Inflation <- GrowthRate(data_us$cpi_all_items)


```




```{r}

for(i in seq_along(data_us)){
  data_us[[i]]<- round(data_us[[i]],3)
}

head(data_us)




for(i in seq_along(data_us[,c(8,12,13)])){
  data_us[,c(8,12,13)][[i]] <- round(log(data_us[,c(8,12,13)][[i]]),3)
}
head(data_us)

 
```



```{r}
#data_us <- data_us %>% select(DATE, quarterlygdpgro, quarterlyrealgd, everything())

#data_dep <- data_us %>% select(DATE, quarterlygdpgro, inflation) 
#data_ind <- data_us%>% select(!c(quarterlygdpgro,inflation,DATE))
  

#data_master <- cbind(data_dep, data_ind)
#head(data_master)

```


dealing with outliers
```{r}
summary(data_us)


#gdp_bar <- quantile(data_master$quarterlygdpgro, 0.98)

#inflation_bar <- quantile(data_master$inflation, 0.99)


#rescaled <- function(x){  ifelse(x>gdp_bar,gdp_bar,x)}


#for (i in seq_along(data_master$quarterlygdpgro)){
#  data_master$quarterlygdpgro[[i]] <- rescaled(data_master$quarterlygdpgro[[i]])
#}


#data_master[-1]%>%map_dbl(median)

```


Correlation matrix
```{r, fig.align='center', fig.width=7, fig.height=7}

corr_us <- cor(data_us[,2:13])


ggcorrplot(corr_us,lab = TRUE,type = "upper")
?ggcorrplot
```


```{r fig.align="center", fig.width = 10}

data_pivot <- data_us %>% pivot_longer(-Date, names_to = "variables", values_to="value")

ggplot(data_pivot, aes(Date,value)) + geom_boxplot() +facet_wrap("variables", scale= "free")

```




```{r fig.align="center", fig.width = 10}

Infl <- data_us%>% select(Date, cpi_all_items)

growth <- function(x){
    (x - dplyr::lag(x))/ dplyr::lag(x)
}


Infl <- Infl %>% mutate(Inflation = growth(cpi_all_items))

plot(Inflation, type="l")

ggplot(data_pivot, aes(Date,value)) + geom_line() +facet_wrap("variables", scale= "free")
```



OLS
```{r}
data_us <- data_us %>% select(!c(Quarterly_real_gdp,
                                 Quarterly_gdp_growth_rate))

data_us1 <- data_us %>% select(!Date)

mod1 <- lm(Gdp_per_capita ~ ., data=data_us1)

summary(mod1)


```




```{r}
library(car)
vif(mod1) # Variance 
?vif

```


```{r}
mod2 <- lm(Gdp_per_capita ~ Unrate+ Inflation+ Monthly_home_supply+
             Consumer_conf_index , data = data_us )
summary(mod2)

#acf(mod2$residuals)
```




```{r}
vif(mod2)


coeftest(mod2, vcov = vcovHC)
acf(mod2$residuals)
```


#################################################################
### Train and test set
##################################################

```{r}
dim(data_us)[1] * 0.7 

test_data <- data_us %>% filter(Date >= "2016-04-01")
train_data <- data_us %>% filter(Date < "2016-04-01")

```




## Linear Model  (OLS)
```{r}

mod_lm <- lm(Gdp_per_capita ~ Unrate+ Inflation+ Monthly_home_supply+
             Consumer_conf_index , data = train_data )
summary(mod_lm)

```

#### Training and test error
```{r}
#MSE_lm <-mean((train_data$Gdp_per_capita - mod_lm$fitted.values)^2)
#MSE_lm

MSE_train <- mean(mod_lm$residuals^2)


mod_pred <- predict(mod_lm, newdata = test_data)

MSE_test <- mean((test_data$Gdp_per_capita - mod_pred)^2)


data.frame(cbind(MSE_train, MSE_test))

```




Non-linear model (Regression Splines)
```{r}

library(splines)
library(gam)

#mod_cs <- lm(Gdp_per_capita ~ ns(Unrate,4)+ ns(Inflation,4)+
#               ns(Monthly_home_supply,4)+
#             ns(Consumer_conf_index,4) , data = data_us )




mod_ss <- gam(Gdp_per_capita ~ s(Unrate)+ s(Inflation)+
               s(Monthly_home_supply)+
             s(Consumer_conf_index),  data = train_data)

names(mod_ss)

plot.Gam(mod_ss, col = "blue", se=TRUE)

MSE_ss <- mean(mod_ss$residuals^2)
MSE_ss  ## Train error

pred_ss <- predict(mod_ss, newdata = test_data)

MSE_ss_test <- mean((test_data$Gdp_per_capita - pred_ss)^2)
MSE_ss_test  ### Test error

erro <- data.frame(cbind(MSE_ss,MSE_ss_test))
colnames(erro) <- c("MSE train data", "MSE test data")

```































